---
title: "Geog533 Lab 8"
author: "Alyssa James"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}
SSM <- SST - SSR
SSR <- 1.7
SST <- 2.3
m1 <- SSM/1
m2 <- SSR/22
F <- m1/m2
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|      `r SSM`    |         `r 1`      |       `r m1`   |      `r F`      |
| Residual |       1.7       |        `r 22`      |       `r m2`   |                 |
| Total    |       2.3       |        `r 23`      |                |                 |


### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
a <- 0.46
b <- 0.19
x <- 50000
y <- a*x + b
y
```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?

For each increase in weekly shopping trips, the predicted annual income will increase by $1900. 

### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}
Fcrit <- qf(0.95, df1 = 1, df2 = 22)
if(F > Fcrit){
print("We cannot reject the null hypothesis")
} else {
print("We reject the null hypothesis")}
```



### Question 1(e)
What is the value of the correlation coefficient?
```{r}
r2 <- SSM/SST
r <- sqrt(SSM/SST)
r
```





## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).
```{r}
snowfall <- c(36, 78, 11, 45)
elevation <- c(400, 800, 200, 675)

m <- lm(elevation ~ snowfall)
summary(m)
print("Intercept <- 127.412")
print("Slope Coefficient <- 9.208")
```


### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
Rse <- 192.1
Rse
```


### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r}
qf(0.95, df1 = 1, df2 = 6)
print("We reject the null hypothesis")
```


### Question 2(d)
Find the value of $r^2$.
```{r}
r2 <- 453628/221520
r2
```


### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}
newdf <- data.frame(snowfall, elevation)
newdf$predicted <- m$fitted.values
newdf$residuals <- m$residuals
newdf
```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
anova(m)
Sm <- 453628
Sr <- 221520
St <- 675148
df1 <- 1
df2 <- 2
m1 <- SSm/df1
m2 <- SSr/df2
F2 <- m1/m2
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|     `r Sm`      |        `r df1`     |    `r m1`      |      `r F2`     |
| Residual |     `r Sr`      |        `r df2`     |    `r m2`      |                 |
| Total    |     `r St`      |                    |                |                 |

### Question 2(g)
Graph the data and the regression line.
```{r}
x <- snowfall
y <- elevation
df <- data.frame(x, y)
plot(y~x, data = df)
abline(m,col="red",lwd=2)
segments(x,fitted(m),x,y,col="blue",lty = "dashed")
```





## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
df <- read.csv(file = "Milwaukee_Sales.csv", header = TRUE)
x <- df$SalePrice
y <- df$LotSize
m <- lm(y~x)
summary(m)
```



## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
df <- read.csv(file = "UK_Housing.csv", header = TRUE)
x <- df$price
y <- df$bedrooms
m <- lm(y~x)
summary(m)
```


### Question 4(b) 
(b) repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
x <- df$price
y <- df$bathrooms
m <- lm(y~x)
summary(m)
qf(0.95, df1 = 1, df2 = 497)
print("We reject the null hypothesis")
```


## Question 5
Download the [hometown.csv](http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r}
df <- read.csv(file = "hometown.csv", header = TRUE)
if(!require(leaflet)) {install.packages("leaflet")}
m <- leaflet(width = "100%")
addTiles(m)  
addMarkers(m, lng=-73.737145, lat=40.655969, popup="The birthplace of Alyssa James")
```

