---
title: "Geog533 Lab 10"
author: "Alyssa James"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)
```{r}
library(UScensus2010county)
```

### Question 1(a)
Plot a map of New York counties using the plot function.
```{r}
data(new_york.county10)
shp <- new_york.county10
df <- shp@data
df
plot(shp)
```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r}
install.packages("tmap")
library(tmap)
qtm(shp)
```


### Question 1(c)	
How many counties in New York State?
```{r}
df
print("There are 62 counties in New York State")
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
print("36007")
```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r}
nypop <- df$P0010001
t <- sum(nypop)
t
a <- summary(nypop)
a
library(moments)
b <- skewness(nypop)
b
```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
a <- hist(nypop)
a
b <- boxplot(nypop)
b
```

## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)
```{r}
library(UScensus2010)
if(!require(UScensus2010tract)) install.tract("osx")
library(UScensus2010tract)
```

### Question 2(a)	
Plot a map of New York census tracts using the plot function.
```{r}
data(new_york.tract10)
shp <- new_york.tract10
df <- shp@data
df
plot(shp)
```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
df
nypop <- df$P0010001
x <- sum(nypop)
x
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
BroomeCounty <- county(fips="007",name="Broome",state="NY",level="tract")
plot(BroomeCounty)
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
Broomepop <- sum(BroomeCounty$P0010001)
Broomepop
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
hist(BroomeCounty$P0010001)

boxplot(BroomeCounty$P0010001)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r}
pr <- data.frame(BroomeCounty[,1:5])
pr$ratio <- BroomeCounty$P0010001/200600
pr
```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)
```{r}
library(raster)
library(ncdf4)
```

### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
ndvi <- raster("NDVI.nc")
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
ndvi.rb <- brick("NDVI.nc")
ndvi.rb
```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}
mean(ndvi.rb)
writeRaster(mean(ndvi.rb),filename = "mean_ndvi.tif", overwrite=TRUE)
max(ndvi.rb)
writeRaster(max(ndvi.rb),filename = "max_ndvi.tif", overwrite=TRUE)
```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r}
ndvi2001 <- ndvi.rb[[13:24]]
plot(ndvi2001)
```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r}
hist(ndvi2001)
```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r}
plot(ndvi.rb,7) 
values <- ndvi.rb[150, 50]
values <- click(ndvi.rb, n=1, xy=TRUE)
values <- click(ndvi.rb, n=1, xy=FALSE)
```

